---
title: "Municipal population CSV missing population counts"
from: markdown+emoji
code-annotations: hover
date: "2024-07-16"
date-modified: "2024-07-16"
categories:
    - tabular data
    - missing data
    - "importance: low"
issue-importance: low
status: "active"
---


---

**Status**: {{< meta status >}}

**Summary**: The dataset contains missing population counts for 49 municipalities in the CSV file `poblacion_municipios.csv` in the `zonificacion_municipios` folder, as well as in the `poblacion.csv` file in the `zonificacion` folder.

**Expected Results**: The population data for all municipalities should be available in the CSV file `poblacion_municipios.csv` in the `zonificacion_municipios` folder and/or in the `poblacion.csv` file in the `zonificacion` folder.

*This repository can be opened in a web browser using Binder to explore all issues (under `issues` subfolder) interactively with `R/Rstudio`. Push the button >> [![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/e-kotov/mitma-data-issues/HEAD?urlpath=rstudio){target="_blank"}*

### Steps to Reproduce

1. **Load Data**

Load libraries and define data files.

```{r}
#| output: false

library(tidyverse)
library(sf)
library(here)
library(DT)


municipal_boundaries_data_file <- here("data/raw_data/v2/zonificacion/zonificacion_municipios/zonificacion_municipios.shp")
municipality_names_file <- here("data/raw_data/v2/zonificacion/zonificacion_municipios/nombres_municipios.csv")
municipality_population_file <- here("data/raw_data/v2/zonificacion/zonificacion_municipios/poblacion_municipios.csv")
all_population_file <- here("data/raw_data/v2/zonificacion/poblacion.csv")
```

Load the data and join the municipality names to the boundaries, as well as population from all population file and municipality population file.


```{r}
municipality_boundaries <- read_sf(municipal_boundaries_data_file)
municipality_boundaries_spain_only <- municipality_boundaries |> 
    filter(! grepl("FR|PT|externo", ID) )

municipality_names <- read_delim(municipality_names_file,
    delim = "|", show_col_types = FALSE, name_repair = "unique_quiet")

municipality_population <- read_delim(municipality_population_file, col_names = c("ID", "population"),
    delim = "|", show_col_types = FALSE, name_repair = "unique_quiet")

all_population <- read_delim(all_population_file,
    delim = "|", show_col_types = FALSE, name_repair = "unique_quiet")

municipality_boundaries_spain_only <- municipality_boundaries_spain_only |>
    left_join(municipality_names |> select(ID, name), by = c("ID")) |> 
    left_join(municipality_population, by = c("ID")) |> 
    left_join(all_population |>
                  group_by(municipio) |>
                  summarise(population_all = sum(poblacion, na.rm = TRUE), .groups = "drop") |> 
                  rename(ID = municipio),
              by = c("ID"))
```

### Results

2. **Missing population (loaded from the `poblacion_municipios.csv` in `zonificacion_municipios` folder)**


```{r}
municipality_boundaries_spain_only |>
    filter(is.na(population)) |>
    nrow()
```


3. **Population data for municipalities also unavailable in the `poblacion.csv` in the `zonificacion` **

There are no population counts for these districts in the `poblacion.csv` file either.

```{r}
municipality_boundaries_spain_only |>
    filter(!is.na(population_all)) |>
    filter(population_all == 0) |> 
    nrow()
```


4. **Names of Municipalities with missing population data**

```{r}
DT::datatable(municipality_boundaries_spain_only |>
                  st_drop_geometry() |> 
                  filter(is.na(population))
                  )
```



### Links to the original files

```{r}
source(here("R/901-download-helpers.R"))
files <- load_latest_v2_xml()

# Filter relevant files
relevant_files <- files |> 
  filter(basename(local_path) %in% basename(c(municipal_boundaries_data_file, municipality_names_file,
                                              municipality_population_file, all_population_file)) )

# Create HTML links
relevant_files <- relevant_files |> 
  mutate(target_url = paste0("<a href='", target_url, "' target='_blank'>", target_url, "</a>"))

# Render the DT table with links
datatable(relevant_files, escape = FALSE, options = list(pageLength = 5))
```


