---
title: "Invalid Geometries in Large Urban Areas Boundaries"
from: markdown+emoji
code-annotations: hover
date: "2024-07-17"
date-modified: "2024-08-17"
categories:
    - spatial data
    - geometry
    - "importance: medium"
issue-importance: "2 - medium"
status: "✅ resolved"
---

---

**Status**: {{< meta status >}}

**Summary**: There are 3 invalid geometries in the Large Urban Areas boundaries dataset.

**Expected Results**: The dataset should not contain any invalid geometries.

---

**Resolution**: resolved with [https://github.com/Robinlovelace/spanishoddata/pull/25](https://github.com/Robinlovelace/spanishoddata/pull/25){target='_blank'} in [`{spanishoddata}`](https://github.com/Robinlovelace/spanishoddata){target='_blank'} R package. 

To get valid spatial data, use `spanishoddata::`[`spod_get_zones()`](https://robinlovelace.github.io/spanishoddata/reference/spod_get_zones.html){target='_blank'}

```{r}
#| eval: false
install.packages("spanishoddata",
  repos = c("https://e-kotov.r-universe.dev", "https://cloud.r-project.org"))

zones <- spanishoddata::spod_get_zones("gau", ver = 2)
```


---

### Steps to Reproduce


1. **Load Data**

Load libraries and define data files.

```{r}
#| output: false

library(tidyverse)
library(sf)
library(here)
library(DT)


gau_boundaries_data_file <- here("data/raw_data/v2/zonificacion/zonificacion_GAU/zonificacion_gaus.shp")
gau_names_file <- here("data/raw_data/v2/zonificacion/zonificacion_GAU/nombres_gaus.csv")
```

Load the data and join the Large Urban Areas names to the boundaries.


```{r}
gau_boundaries <- read_sf(gau_boundaries_data_file)

gau_names <- read_delim(gau_names_file,
    delim = "|", show_col_types = FALSE, name_repair = "unique_quiet")

gau_boundaries <- gau_boundaries |>
    left_join(gau_names |> select(ID, name), by = c("ID"))
```


2. **Identify Invalid Geometries**

```{r}
gau_boundaries_invalid <- !st_is_valid(gau_boundaries)
```

### Results

3. **Number of Invalid Geometries**

```{r}
sum(gau_boundaries_invalid)
```

4. **Names of Large Urban Areas with Invalid Geometries**

```{r}
DT::datatable(gau_boundaries[gau_boundaries_invalid,] |> st_drop_geometry())
```



### Links to the original files

```{r}
source(here("R/901-download-helpers.R"))
files <- load_latest_v2_xml()

# Filter relevant files
relevant_files <- files |> 
  filter(basename(local_path) %in% basename(c(gau_boundaries_data_file, gau_names_file)) )

# Create HTML links
relevant_files <- relevant_files |> 
  mutate(target_url = paste0("<a href='", target_url, "' target='_blank'>", target_url, "</a>"))

# Render the DT table with links
datatable(relevant_files, escape = FALSE, options = list(pageLength = 5))
```
