---
title: "Duplicate and missing municipality data"
from: markdown+emoji
code-annotations: hover
date: "2024-08-17"
date-modified: "2024-08-17"
categories:
    - tabular data
    - incorrect data
    - zone IDs mismatch
    - missing data
    - "importance: high"
issue-importance: "3 - high"
status: "⚠️ active"
---

---

**Status**: {{< meta status >}}

**Importance**: {{< meta issue-importance >}}

**Summary**: Folders for municipality data contain files with disctrict data.

**Expected Results**: There should be no district data files in municipality folders. There should be no missig municipality data.

---


---


### Steps to Reproduce

1. **Load Data**

Load libraries and define data files.

```{r}
#| output: false

library(tidyverse)
library(sf)
library(here)
library(DT)
library(spanishoddata)

Sys.setenv("SPANISH_OD_DATA_DIR" = here::here("data"))
v1 <- spanishoddata:::spod_available_data(1)
v1$local_path <- fs::path_rel(v1$local_path)
```

Folders with municipality data contain files with disctrict data.

```{r}
DT::datatable(v1 |> dplyr::filter(grepl("municipios.*distrito", target_url)))
```

There is no issue with municipality data in district folders.

```{r}
DT::datatable(v1 |> dplyr::filter(grepl("distrito.*municipios", target_url)))
```

Is the data in the municipality folders that is labelled as district actually district level and are there files for municipalities in the same folders?

```{r}
problematic_urls <- v1 |> dplyr::filter(grepl("municipios.*distrito", target_url)) |> pull(target_url)
problematic_dates <- str_extract(problematic_urls, "[0-9]{8}") |> paste0(collapse = "|")
DT::datatable(v1 |> dplyr::filter(grepl(problematic_dates, target_url)) |> dplyr::arrange(local_path))
```

### What we find

- `20210205_maestra_2_mitma_distrito.txt.gz` in `maestra2-mitma-municipios/ficheros-diarios/2021-02/` folder should not be used in any batch imports, as there is `20210205_maestra_2_mitma_municipio.txt.gz` in the same folder.

- Critically, only `20200712_maestra_1_mitma_distrito.txt.gz` is in municipalities folder, there is no file `20200712_maestra_1_mitma_municipio.txt.gz` in the same folder. This file is also missing from the tar archive with monthly data ([https://opendata-movilidad.mitma.es/maestra1-mitma-municipios/meses-completos/202007_maestra1_mitma_municipio.tar](https://opendata-movilidad.mitma.es/maestra1-mitma-municipios/meses-completos/202007_maestra1_mitma_municipio.tar)).

### Conclusion and how the issue should be fixed

- The district files placed in the municipality folders should not be donwloaded to avoid data contamination during batch imports. That concerns files `20210205_maestra_2_mitma_distrito.txt.gz` in `maestra2-mitma-municipios/ficheros-diarios/2021-02/`.

- `20200712_maestra_1_mitma_distrito.txt.gz` may be aggregated to municipality level using the reference tables:

```{r}
relaciones_distrito_mitma <- readr::read_delim(
  "https://opendata-movilidad.mitma.es/relaciones_distrito_mitma.csv", delim = "|", show_col_types = FALSE)
DT::datatable(relaciones_distrito_mitma)
```

```{r}
relaciones_municipio_mitma <- readr::read_delim(
  "https://opendata-movilidad.mitma.es/relaciones_municipio_mitma.csv", delim = "|", show_col_types = FALSE)
DT::datatable(relaciones_municipio_mitma)
```

However, the issue is that [https://opendata-movilidad.mitma.es/maestra1-mitma-distritos/ficheros-diarios/2020-07/20200712_maestra_1_mitma_distrito.txt.gz](https://opendata-movilidad.mitma.es/maestra1-mitma-distritos/ficheros-diarios/2020-07/20200712_maestra_1_mitma_distrito.txt.gz) (has 5237350 rows) is not the same as "[https://opendata-movilidad.mitma.es/maestra1-mitma-municipios/ficheros-diarios/2020-07/20200712_maestra_1_mitma_distrito.txt.gz](https://opendata-movilidad.mitma.es/maestra1-mitma-municipios/ficheros-diarios/2020-07/20200712_maestra_1_mitma_distrito.txt.gz) (has 5020279), while the latter is also clearly not a municipality file in disguise, as it is noticeable larger and mostly contains district IDs, not municipality IDs. It is probably safer to download the larger file.
